---
title: "Chase_Twitter"
author: "Anurag Garg"
date: "November 15, 2015"
output: pdf_document
---

##1. Getting Data

To get data for this project, we need to authorize first and then we can access twitter API. Following script will do it and the output will be a data frame of 3200 tweets for @Chase. 
```{r,message=FALSE,eval=FALSE}
#install_github("geoffjentry/twitteR")

# Initialization of required packages
library(RCurl)
library(RJSONIO)
library(digest)
library(devtools)
library(twitteR)

# Download the certificate needed for authentication
 if (!file.exists('cacert.perm')){
   download.file(url = 'http://curl.haxx.se/ca/cacert.pem',
                 destfile='cacert.perm')
 }

# Setup twitter keys and secrets

# These can be accessed from your application's page on dev.twitter.com
# under the names Consumer key, Consumer secret, Access token, and
# Access token secret respectively.

consumerKey = '8jqqM7OlEk6Fkdnc0aGLjUtXM'
consumerSecret = '8U61tPinhxbGRYJnCk257RUPNT0umxybwN0HVqrQwxuH3sfmEk'
accessToken = '2198592541-R6BnXREHUn6SeMVI9w1fZPof4jpVnr41p1RNJ92'
accessTokenSecret = 'aRp5f97FdrT9dXYTMZDtzw6CQPfHaLlHYNg3r79jOzkPs'
setup_twitter_oauth(consumerKey, consumerSecret, 
                    accessToken, accessTokenSecret)

#get the tweets for @Chase.

tweets<-searchTwitter("@Chase", n=3200)
#length(tweets)

# Transform tweets list into a data frame
tweets.df = do.call("rbind",lapply(tweets,as.data.frame))

# Save data into an R file for using it later.
saveRDS(tweets.df,file="tweets_data.Rda")
```

**Note:**

For most of the tweets (>99%), location (latitude and longitude) is disabled by the user and hence we are not doing any analysis related to location.


##2. Cleaning Data

Now we will perform the task of cleaning data. Data need to be cleaned because:

a. We will remove the columns which are not needed. We will retain following 5 columns for our analysis: 
      + text
      + created
      + id
      + screenName
      + retweetCount

b. There are duplicate tweets in the data.

c. We do not need tweets which are tweeted by Chase itself (tweets by the official handle of Chase.)

```{r,message=FALSE}

library("plyr")

# Load the tweets data.

tweets.df<-readRDS(file="tweets_data.Rda")

#clean data text

#1. removing column which are not needed
col_needed<-c("text","created","id","screenName","retweetCount")
needed_col_df<-tweets.df[,col_needed]


#2. removing rows which are duplicate tweets
needed_data_df<- subset(needed_col_df, !duplicated(needed_col_df[,1]))

#3. removing rows where user is chase
data_df<- subset(needed_data_df,!(needed_data_df$screenName=="Chase"))

```

##3. Exploratory Data Analysis

###a. General sentiment about @Chase.
```{r, message=FALSE }

source("ts_score_sentiment.R")
library("plyr")
library(stringr)

#Getting a score for each tweet using sentiment analysis 

# Getting lexicon of positive and negative words.

pos <- scan('positive-words.txt', what='character', comment.char=';')
neg <- scan('negative-words.txt', what='character', comment.char=';')

#Giving a score to a tweet based on net (positive-negative) words.

analysis <- ts_score_sentiment(data_df$text, pos, neg)
data_df<-merge(data_df,analysis)

#Multiplying the score with the retweet count to get the effective score.
#add 1 to retweet count to get effective number of times the tweet is posted.

data_df$effScore<-(data_df$score*(data_df$retweetCount+1))

#Plotting a histogram to see score distribution for chase at this point.
hist(data_df$score, probability = TRUE,xlab = "Tweet score",
     main = "Distribution of tweet scores")

```

We can see that density distribution of the tweet scores is nearly normal and also the histogram is little misleading. Although maximum tweets are in range of 0 to -1 which shows a little negative outlook but from the normal distribution we can see that the mean score is 0 which means a neutral outlook.


###b. Sentiment about individual products

We will now categorize the tweets according to different products offered by Chase. Major products which we considered are:

i. Sapphire Credit Card

ii. Freedom Credit Card

iii. Slate Credit Card

iv. Credit Card(Misc tweets about credit cards)

v. Checking account

vi. Debit card


```{r, message=FALSE }

#categorize data according to product

#Sapphire
data_sapphire<- subset(data_df,grepl("sapphire",data_df$text,
                                                   ignore.case = T))
data_sapphire$product<-rep("Sapphire",nrow(data_sapphire))

#Freedom
data_freedom<- subset(data_df,grepl("freedom",data_df$text,
                                                   ignore.case = T))
data_freedom$product<-rep("Freedom",nrow(data_freedom))

#Slate
data_slate<- subset(data_df,grepl("slate",data_df$text,
                                                   ignore.case = T))      
data_slate$product<-rep("Slate",nrow(data_slate))

#Misc Credit card
data_allcredit<-rbind.data.frame(data_slate, data_freedom,data_sapphire)
data_credit<- subset(data_df,!(data_df$id %in% data_allcredit$id))
credit_match<-c("credit card","creditcard")
data_credit<- subset(data_credit,grepl(paste(credit_match,collapse="|"),data_credit$text,
                                                ignore.case = T))      
data_credit$product<-rep("Misc_Credit",nrow(data_credit))

#Checking Account
checking_match<-c("checking account","checkingaccount","checking")
data_checkingacc<- subset(data_df,grepl(paste(checking_match,collapse="|"),data_df$text,
                                                ignore.case = T))  
data_checkingacc$product<-rep("Checking",nrow(data_checkingacc))

#Debit card
debit_match<-c("debit card","debitcard")
data_debit<- subset(data_df,grepl(paste(debit_match,collapse="|"),data_df$text,
                                                ignore.case = T))   
data_debit$product<-rep("Debit",nrow(data_debit))

```

###i. Sentiment analysis of Sapphire credit card.

```{r,echo=FALSE,message=FALSE}

#Plotting a histogram to see general viewpoint of Sapphire credit card.

hist(data_sapphire$effScore, probability = FALSE,xlab = "Tweet score", 
  main = "Distribution of tweet scores for Sapphire")

```

###ii. Sentiment analysis of Freedom credit card.

```{r,echo=FALSE,message=FALSE}

#Plotting a histogram to see general viewpoint of Freedom credit card.

hist(data_freedom$effScore, probability = F,xlab = "Tweet score", 
    main = "Distribution of tweet scores for Freedom")

```

###iii. Sentiment analysis of Slate credit card.

```{r,echo=FALSE,message=FALSE}

#Plotting a histogram to see general viewpoint of Slate credit card.

hist(data_slate$effScore, probability = FALSE,xlab = "Tweet score", 
    main = "Distribution of tweet scores for Slate")

```

###iv. Sentiment analysis of Misc. tweets about credit card.

```{r,echo=FALSE,message=FALSE}

#Plotting a histogram to see general viewpoint of Misc. credit card.

hist(data_credit$effScore, probability = FALSE,xlab = "Tweet score", 
    main = "Distribution of tweet scores for Misc. credit card")

```

###v. Sentiment analysis of Checking account.

```{r,echo=FALSE,message=FALSE}

#Plotting a histogram to see general viewpoint of Checking account.

hist(data_checkingacc$effScore, probability = FALSE,xlab = "Tweet score", 
    main = "Distribution of tweet scores for Checking account")

```

###vi. Sentiment analysis of Debit card.

```{r,echo=FALSE,message=FALSE}

#Plotting a histogram to see general viewpoint of Debit card.

hist(data_debit$effScore, probability = FALSE,xlab = "Tweet score", 
     main = "Distribution of tweet scores for Debit card")

```

We can see from histograms that products with positive outlook are:
  
  + Freedom
  + Misc. credit cards

  
We can see from histograms that products with negative outlook are:
  
  + Sapphire
  + Debit card
  
  
**Note:**

For **Slate credit card** and **Checking account**, We can not say anything because the number of observations are pretty low.

##4. Graph Analysis

We can create a bi-partite graph of users and products. Edges weight can be the score of the tweet which user posted about a specific product.

```{r,echo=FALSE,message=FALSE}
#Merge all the tweets related to products in one data frame.
data_graph<-rbind.data.frame(data_sapphire, data_freedom, data_slate, data_credit,
                             data_checkingacc, data_debit)


```
